Resolvable Ambiguity

### 2019-07-11 paper/paper.tex
Be more positive!

\noindent Unfortunately, this automaton recognizes the word $(((e)))$, which is not the unique linear encoding we require; it should be $(e)$.

@@

@@

\paragraph{Limitations of Syncons}
%
Other than syntactic constructs that are not yet specified, there
are two sources of failure in the parts of the OCaml compiler that
we cannot parse.
%
First, our parsing tool does not support specifying ``longest
match'' on a production, which is required to handle the pattern
matching constructs correctly. Some cases can be handled using
\syncon{forbid} declarations, but it does not get us all the way.
%
Second, our system uses a different definition of precedence than
the OCaml language. Our translation from precedence to implicit
\syncon{forbid} declarations is shallow (it only considers direct
children), while the OCaml has deep precedence. For example,
addition binds stronger than \ocaml{let} (which syntactically
functions as a prefix operator in OCaml), and thus
%
``\ocaml{1 + let x = 1 in x + 2}'' should be parsed as
%
``\ocaml{1 + (let x = 1 in (x + 2))}''. Our tool, however, reports
the expression as ambiguous, additionally suggesting the
alternative interpretation
%
``\ocaml{(1 + (let x = 1 in x)) + 2}''.
%
This interpretation is indeed valid if we only look at the direct
children of each operator, since ``\ocaml{(_ + _) + 2}'',
``\ocaml{1 + let _ in _}'', and ``\ocaml{let x = 1 in x}'' are all
individually correct with regards to precedence.

>>>>

\noindent Unfortunately, this automaton recognizes the word $(((e)))$,
which is not the unique linear encoding we require; it should be
$(e)$. The reason for this mismatch is that each of the productions
with labels $a$ and $b$ introduces an additional pair of parentheses
in the word recognized by the automaton. To ensure that the automaton
instead recognizes the required unique linear encoding $(e)$, we
extend the construction of the automaton such that ``passing through''
a non-terminal without recognizing any terminals does \emph{not} add
parentheses. The idea is to skip the DFAs of such non-terminals, but
still keep track of the skipped production(s) in the symbols pushed to
the stack; thus, runs which differ in skipped non-terminals are still
distinguished.

For our example, the resulting automaton looks as follows:

@@

Note that the transition from state 8 to state 6 pushes the symbol
$(8, 9, ab)$ to the stack. This means (a) that we started the
``replacement'' for the transition from 8 to 9, and (b) that we passed
through the productions with labels $a$ and $b$ without consuming
terminals, thus ending up in the starting state of the DFA
corresponding to production $c$.

@@

\paragraph{Current Limitations of Syncons}
%
Other than syntactic constructs that are not yet specified, there
are two sources of failure in the parts of the OCaml compiler that
we cannot parse.
%
First, our parsing tool does not support specifying ``longest
match'' on a production, which is required to handle the pattern
matching constructs correctly. Some cases can be handled using
\syncon{forbid} declarations, but it does not get us all the way.
%
Second, our system uses a different definition of precedence than
the OCaml language. Our translation from precedence to implicit
\syncon{forbid} declarations is shallow (it only considers direct
children), while the OCaml has deep precedence. For example,
addition binds stronger than \ocaml{let} (which syntactically
functions as a prefix operator in OCaml), and thus
%
``\ocaml{1 + let x = 1 in x + 2}'' should be parsed as
%
``\ocaml{1 + (let x = 1 in (x + 2))}''. Our tool, however, reports
the expression as ambiguous, additionally suggesting the
alternative interpretation
%
``\ocaml{(1 + (let x = 1 in x)) + 2}''.
%
This interpretation is indeed valid if we only look at the direct
children of each operator, since ``\ocaml{(_ + _) + 2}'',
``\ocaml{1 + let _ in _}'', and ``\ocaml{let x = 1 in x}'' are all
individually correct with regards to precedence.
